---
title: "pstat 131 hw 3"
output: html_document
---

```{r, message= FALSE}
library(tidymodels)
library(tidyverse)
library(ISLR) 
library(ISLR2) 
library(discrim)
library(poissonreg)
library(corrr)
library(corrplot)
library(klaR) 
library(pROC)
tidymodels_prefer()
```
```{r}
titanic<- read.csv('/Users/ankitapattnaik/Downloads/homework-3/data/titanic.csv', stringsAsFactor=T)
titanic.cdbk<- read_csv('/Users/ankitapattnaik/Downloads/homework-3/data/titanic_codebook.txt')
```


1. Split the data, stratifying on the outcome variable, survived. You should choose the proportions to split the data into. Verify that the training and testing data sets have the appropriate number of observations. Take a look at the training data and note any potential issues, such as missing data.

```{r}
set.seed(3435)

titanic_split <- initial_split(titanic, prop = 0.70,
                                strata = survived)
titanic_train <- training(titanic_split)
titanic_test <- testing(titanic_split)

titanic_train
```

Why is it a good idea to use stratified sampling for this data?
- It is a good idea to use stratified sampling for this data because by taking a part of the data, we can make assumptions based on demographics and draw conclusions on all the data based off the stratified chunk. This allows us to work with less data and be more clean with our work, that way we can draw conclusions on the population as a whole. 




2. Using the training data set, explore/describe the distribution of the outcome variable survived.

```{r}
titanic_train %>% 
  ggplot(aes(x = survived)) +
  geom_bar()
```

- Less people survived than not. We see this through a bar plot. 



3. Using the training data set, create a correlation matrix of all continuous variables. Create a visualization of the matrix, and describe any patterns you see. 

```{r}

cor_titanic <- titanic_train %>%
  select(where(is.numeric)) %>%
  correlate()
rplot(cor_titanic) 

```

Are any predictors correlated with each other? Which ones, and in which direction?
- It seems that sib_sp (# of spouses and siblings abroad the titanic) and parch (# of children and parents on the titanic) have strong positive correlation. 
- fare and parch have distinct positive correlation.
- sib_sp and fare, age and fare, sib_sp and pclass all have very little positive correlation
- pclass (ticket class) and fare (passenger fare) have a strong negative correlation. 
- age and pclass, age and sib_sp have distinct negative correlation
- age and parch have light negative correlation
- passenger_id and sib_sp have light negative correlation. 



4. Using the training data, create a recipe predicting the outcome variable survived. Include the following predictors: ticket class, sex, age, number of siblings or spouses aboard, number of parents or children aboard, and passenger fare.

Next, use step_dummy() to dummy encode categorical predictors. Finally, include interactions between:

Sex and passenger fare, and
Age and passenger fare.


```{r}
titanic_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + 
                           parch + fare, data = titanic_train) %>%
  step_impute_linear(age) %>%
  step_dummy(all_nominal_predictors()) %>% 
  step_interact(terms = ~ starts_with("sex"):fare) %>%
  step_interact( ~ age:fare) 
```



5. Specify a logistic regression model for classification using the "glm" engine. Then create a workflow. Add your model and the appropriate recipe. Finally, use fit() to apply your workflow to the training data.

```{r}
log_reg <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

log_wkflow <- workflow() %>% 
  add_model(log_reg) %>% 
  add_recipe(titanic_recipe)

log_fit <- fit(log_wkflow, titanic_train)
```



6. Repeat Question 5, but this time specify a linear discriminant analysis model for classification using the "MASS" engine.

```{r}
lda_mod <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")

lda_wkflow <- workflow() %>% 
  add_model(lda_mod) %>% 
  add_recipe(titanic_recipe)

lda_fit <- fit(lda_wkflow, titanic_train)
```



7. Repeat Question 5, but this time specify a quadratic discriminant analysis model for classification using the "MASS" engine.

```{r}
qda_mod <- discrim_quad() %>% 
  set_mode("classification") %>% 
  set_engine("MASS")

qda_wkflow <- workflow() %>% 
  add_model(qda_mod) %>% 
  add_recipe(titanic_recipe)

qda_fit <- fit(qda_wkflow, titanic_train)
```



8.Repeat Question 5, but this time specify a naive Bayes model for classification using the "klaR" engine. Set the usekernel argument to FALSE. 

```{r}
nb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

nb_wkflow <- workflow() %>% 
  add_model(nb_mod) %>% 
  add_recipe(titanic_recipe)

nb_fit <- fit(nb_wkflow, titanic_train)
```



9. Now you’ve fit four different models to your training data.

Use predict() and bind_cols() to generate predictions using each of these 4 models and your training data. Then use the accuracy metric to assess the performance of each of the four models.

```{r}
titanic_log<-predict(log_fit, new_data = titanic_train, type = "prob")
titanic_log_col<-bind_cols(titanic_log, titanic_train)
log_reg_acc<- augment(log_fit, new_data= titanic_train) %>% accuracy(truth=survived, estimate=.pred_class)
log_reg_acc

titanic_lda<-predict(lda_fit, new_data = titanic_train, type = "prob")
titanic_lda_col<-bind_cols(titanic_lda, titanic_train)
lda_reg_acc<- augment(lda_fit, new_data= titanic_train) %>% accuracy(truth=survived, estimate=.pred_class)
lda_reg_acc

titanic_qda<-predict(qda_fit, new_data = titanic_train, type = "prob")
titanic_qda_col<-bind_cols(titanic_qda, titanic_train)
qda_reg_acc<- augment(qda_fit, new_data= titanic_train) %>% accuracy(truth=survived, estimate=.pred_class)
qda_reg_acc

titanic_nb<-predict(nb_fit, new_data = titanic_train, type = "prob")
titanic_nb_col<-bind_cols(titanic_nb, titanic_train)
nb_reg_acc<- augment(nb_fit, new_data= titanic_train) %>% accuracy(truth=survived, estimate=.pred_class)
nb_reg_acc

accuracies <- c(log_reg_acc$.estimate, lda_reg_acc$.estimate, 
                qda_reg_acc$.estimate, nb_reg_acc$.estimate)
models <- c("Logistic Regression", "LDA", "Naive Bayes", "QDA")
results <- tibble(accuracies = accuracies, models = models)
results %>% 
  arrange(-accuracies)
```
Which model achieved the highest accuracy on the training data?
- The logistic regression model has the highest accuracy on the training data due to having the highest estimate, followed by LDA, Naive Bayes, and QDA. 



10. Fit the model with the highest training accuracy to the testing data. Report the accuracy of the model on the testing data.

Again using the testing data, create a confusion matrix and visualize it. Plot an ROC curve and calculate the area under it (AUC).

How did the model perform? Compare its training and testing accuracies. If the values differ, why do you think this is so?

```{r}
predict(log_fit, new_data = titanic_test, type = "prob")

augment(log_fit, new_data = titanic_test) %>%
  conf_mat(truth = survived, estimate = .pred_class)

augment(log_fit, new_data = titanic_test) %>%
  roc_curve(survived, .pred_No) %>%
  autoplot()

augment(log_fit, new_data = titanic_test) %>%
  roc_auc(survived, .pred_No) 




```

I would assume the model did well because the area under the curve is 0.88 which represents the probability of the survived not surviving at 88% which is quite high. There is a 0.07 difference between the test and training data accuracies with binary estimator. The values might a little because of the methods of measuring accuracies, as well as the data that correlates with the split data. The test and training data both have a different amount of observations so that is why the observations might predict a slightly different outcome. 


